---
title: "Agreement analysis"
author: "Harry Clark and Matt Nolan"
date: "01/11/2023"
output: html_document
---

```{r}
library(tidyverse)
library(broom)
library(lme4)
library(Matrix)
library(lmerTest)
library(emmeans)
library(glmmTMB)
```

# subset the data acccording to trial type
```{r}

agreement_df <- read_rds(file="agreement_df.Rda")


reformat_agreement <- function(df) {
    df$classifier_i[df$classifier_i != "G"] <- "NG"
    df$classifier_j[df$classifier_j != "G"] <- "NG"
    df$comp <- as_factor(paste(df$classifier_i,df$classifier_j, sep="_"))
    df$agreement <- unlist(df$agreement)
    df$shuffled <- unlist(df$shuffled)
    df$cluster_id_i <- unlist(df$cluster_id_i)
    df$cluster_id_j <- unlist(df$cluster_id_j)

    df
}

agreement_df <- reformat_agreement(agreement_df)

# Make a version of the data frame excluding pairs that are completely identical (agreement = 1) or different (agreement = 0)
agreement_df_ex <- agreement_df[agreement_df$agreement != 0,]
agreement_df_ex <- agreement_df_ex[agreement_df_ex$agreement != 1,]

# squeeze data to modify 0s and 1s
agreement_df$agreement_s <- (agreement_df$agreement*(length(agreement_df$agreement)-1) + 0.5)/length(agreement_df$agreement)
```


# look at the data
```{r}
ggplot(agreement_df[agreement_df$shuffled==0,], aes(comp, agreement)) +
    geom_violin() +
    geom_jitter(alpha = 0.05)
```

# Mixed effect model to compare three groups (experimental data only)
Problem with this is that the data is bounded between 0 and 1, and errors are unlikely to follow a normal distribution.
```{r}
formula_all <- formula(agreement ~ comp + (1 | mouse_id / session_id))

all_mod <- lmerTest::lmer(formula = formula_all, data = agreement_df[agreement_df$shuffled==0,])

# This version takes longer but generated correct df
cons <- emmeans(all_mod, "comp", pbkrtest.limit = 20000, lmerTest.limit=20000)
#cons <- emmeans(all_mod, "comp")
pairs(cons)
pwpm(cons)

plot(cons)
```

# Try using a beta distribution instead to model the data
This can be fit with the glmmTB package

```{r}
model_full <- glmmTMB(agreement_s ~ comp + (1 | mouse_id / session_id), data = agreement_df[agreement_df$shuffled==0,], family = beta_family())


model_null<- glmmTMB(agreement_s ~ (1 | mouse_id / session_id), data = agreement_df[agreement_df$shuffled==0,], family = beta_family())

# compare models
anova(model_full, model_null)

# Use emmeans to evaluate factors
comps <- emmeans(model_full, "comp")
pairs(comps)
pwpm(comps)

plot(comps)
```

Try again but using df with 0s and 1s excluded. 
```{r}
model_full <- glmmTMB(agreement ~ comp + (1 | mouse_id / session_id), data = agreement_df_ex[agreement_df_ex$shuffled==0,], family = beta_family())


model_null<- glmmTMB(agreement ~ (1 | mouse_id / session_id), data = agreement_df_ex[agreement_df_ex$shuffled==0,], family = beta_family())

# compare models
anova(model_full, model_null)

# Use emmeans to evaluate factors
comps <- emmeans(model_full, "comp")
pairs(comps)
pwpm(comps)

plot(comps)
```

This seems potentially problematic. Significance looks 'better' with 1s and 0s excluded but there is no good reason to do this.

# Use differences instead
Try similar analysis but comparing differences for each cell pair between it's agreement score and the average of the corresponding shuffled agreement scores.


```{r}
c_a_df <- read_rds(file="compiled_agreement_df.Rda")

c_a_re_df <- reformat_agreement(c_a_df)

c_a_re_df <- pivot_wider(c_a_re_df, names_from = "shuffled", values_from = "agreement")

c_a_re_df$diff <- c_a_re_df$'0' - c_a_re_df$'1'
```

Plot to replication 4E top.
```{r}
ggplot(c_a_re_df, aes(comp, diff)) +
    geom_violin() +
    geom_jitter(alpha = 0.05)

```

Now fit model to the differences

```{r}
formula_all <- formula(diff ~ comp + (1 | mouse_id / session_id))

all_mod <- lmerTest::lmer(formula = formula_all, data = c_a_re_df)

# This version takes longer but generated correct df
cons <- emmeans(all_mod, "comp", pbkrtest.limit = 20000, lmerTest.limit=20000)
# This version is faster
#cons <- emmeans(all_mod, "comp")
pairs(cons)
# Compact visualisation of comparisons
pwpm(cons)

plot(cons)

```


Or use more complex model
```{r}
formula_all <- formula(diff ~ comp + (1 | mouse_id / session_id / cluster_id_i) + (1 | mouse_id / session_id / cluster_id_j))

all_mod <- lmerTest::lmer(formula = formula_all, data = c_a_re_df)

# This version takes longer but generated correct df
#cons <- emmeans(all_mod, "comp", pbkrtest.limit = 20000, lmerTest.limit=20000)
# This version is faster
cons <- emmeans(all_mod, "comp")
pairs(cons)
# Compact visualisation of comparisons
pwpm(cons)

plot(cons)

```

# Next compare agreement scores between shuffled and data groups
This is for analyses for Figure 4E bottom.

Use data with agreement scores of 1 and 0 removed, 'agreement_df_ex'
Assume delta distributions for agreement scores.
```{r}

mod_fit <- function(df, comp_str = "G_G") {

    mod_data <- df[df$comp==comp_str,]

    model_full <- glmmTMB(agreement ~ shuffled + (1 | mouse_id / session_id), data = mod_data, family = beta_family())

    model_null <- glmmTMB(agreement ~ (1 | mouse_id / session_id), data = mod_data, family = beta_family())
    
    model_comp <- anova(model_full, model_null)

    return(model_comp)
}

model_comp_1 <- mod_fit(df = agreement_df_ex, comp_str =  "G_G")
model_comp_1

model_comp_2 <- mod_fit(df = agreement_df_ex, comp_str = "G_NG")
model_comp_2

model_comp_3 <- mod_fit(df = agreement_df_ex, comp_str = "NG_NG")
model_comp_3
```


Repeat with a model that also accounts for cluster_id. Think of this as cluster_id_j is nested within cluster_id_i, which in turn is nested within session and then within mouse.
```{r}

mod_fit <- function(df, comp_str = "G_G") {

    mod_data <- df[df$comp==comp_str,]

    model_full <- glmmTMB(agreement ~ shuffled + (1 | mouse_id / session_id / cluster_id_i / cluster_id_j), data = mod_data, family = beta_family())

    model_null <- glmmTMB(agreement ~ (1 | mouse_id / session_id / cluster_id_i / cluster_id_j), data = mod_data, family = beta_family())
    
    model_comp <- anova(model_full, model_null)

    return(model_comp)
}

model_comp_1 <- mod_fit(df = agreement_df_ex, comp_str =  "G_G")
model_comp_1

model_comp_2 <- mod_fit(df = agreement_df_ex, comp_str = "G_NG")
model_comp_2

model_comp_3 <- mod_fit(df = agreement_df_ex, comp_str = "NG_NG")
model_comp_3
```

