---
title: "Compare firing properties across session-level firing modes for stable cells"
author: "Matt Nolan"
date: "01/11/2023"
output: html_document
---

# Setup
```{r}
library(tidyverse)
library(lme4)
library(emmeans)
```


# load data
```{r}
cells_data <- readRDS(file = "cells_df_singular_values.Rda")
cells_data <- filter(cells_data, grid_cell == "TRUE")
cells_data$Lomb_classifier_ <- as_factor(cells_data$Lomb_classifier_)
```

# look at data
Replicate Figure 3 Figure Supplement 4.

This is a similar analysis to Figure 2F, but instead of using the session based classification of TIG vs TAG, it uses stability of TIG vs TAG classification within a session.

Thus, whereas in Figure 2F we use the column 'Lomb_classifier_' to group cells, here we use the following:
rolling:proportion_encoding_X, where X is distance, null or position. These variables show how much of the session is encoded by a single mode.

Cells that are "stable" have a proportion above 0.85

First extract classifications for each cell
```{r}
cells_data <- cells_data %>%
    rename(
        prop_TI = 'rolling:proportion_encoding_distance',
        prop_TA = 'rolling:proportion_encoding_position'
    )

thresh <- 0.85

cells_data$stable_cat <- "None"
cells_data$stable_cat[cells_data$prop_TI > thresh] <- "TI"
cells_data$stable_cat[cells_data$prop_TA > thresh] <- "TA"
```

Firing rate
```{r}
ggplot(cells_data, aes(stable_cat, mean_firing_rate_vr)) +
    geom_boxplot()
```

Spatial information (bits / s)
```{r}
ggplot(cells_data, aes(stable_cat, spatial_information_score_Isec_vr)) +
    geom_boxplot()
```

Spatial information (bits / s)
Distribution is more symmetrical with log transformation
```{r}
ggplot(cells_data, aes(stable_cat, log(spatial_information_score_Isec_vr))) +
    geom_boxplot()
```

Spatial information (bits / spike)
This is less informative as doesn't account for differences in firing rate.
```{r}
ggplot(cells_data, aes(stable_cat, spatial_information_score_Ispike_vr)) +
    geom_boxplot()
```



# Evaluate normality
log transformation improves fit
```{r}
qqnorm(log(cells_data$mean_firing_rate_vr), pch = 1, frame = FALSE)
qqline(log(cells_data$mean_firing_rate_vr), col = "steelblue", lwd = 2)
```

```{r}
qqnorm(log(cells_data$spatial_information_score_Isec_vr), pch = 1, frame = FALSE)
qqline(log(cells_data$spatial_information_score_Isec_vr), col = "steelblue", lwd = 2)
```


```{r}
qqnorm(cells_data$ML_SNRs, pch = 1, frame = FALSE)
qqline(cells_data$ML_SNRs, col = "steelblue", lwd = 2)
```

```{r}
qqnorm(log(cells_data$ML_peak_width), pch = 1, frame = FALSE)
qqline(log(cells_data$ML_peak_width), col = "steelblue", lwd = 2)
```



And some summary statistics
```{r}
cells_data %>% group_by(stable_cat) %>% reframe(mean = mean(spatial_information_score_Isec_vr), stdev = sd(spatial_information_score_Isec_vr))

cells_data %>% group_by(stable_cat) %>% reframe(mean = mean(mean_firing_rate_vr), stdev = sd(mean_firing_rate_vr))

cells_data %>% group_by(stable_cat) %>% reframe(mean = mean(ML_SNRs), stdev = sd(ML_SNRs))
```




## Spatial information

# fit model and compare to a null model
Choose a gamma family as the data is continuous and > 0.
```{r}
si_full_mod <- glmer(spatial_information_score_Isec_vr ~ stable_cat + (1 | mouse / session_id), data = cells_data, family = Gamma)

summary(si_full_mod)

si_null_mod <- glmer(spatial_information_score_Isec_vr ~ (1 | mouse / session_id), data = cells_data, family = Gamma)

summary(si_null_mod)

anova(si_full_mod, si_null_mod)
```

Or fit log transformed data with linear mixed effect model.
```{r}
si_full_mod <- lmer(log(spatial_information_score_Isec_vr) ~ stable_cat + (1 | mouse / session_id), data = cells_data)

summary(si_full_mod)

si_null_mod <- lmer(log(spatial_information_score_Isec_vr) ~ (1 | mouse / session_id), data = cells_data)

summary(si_null_mod)

anova(si_full_mod, si_null_mod)
```



# Use emmeans package to evaluate marginal means from the fit model
```{r}
si_emm <- emmeans(si_full_mod, "stable_cat")

pairs(si_emm)

plot(si_emm)
```




## Spike rate
# fit model and compare to a null model
Could choose a gamma family as the data is continuous and > 0
```{r}
fr_full_mod <- glmer(mean_firing_rate_vr ~ stable_cat + (1 | mouse / session_id), data = cells_data, family = Gamma)

summary(fr_full_mod)

fr_null_mod <- glmer(mean_firing_rate_vr ~ (1 | mouse / session_id), data = cells_data, family = Gamma)

summary(fr_null_mod)

anova(fr_full_mod, fr_null_mod)
```

Or instead fit linear mixed effect model to log transformed values
```{r}
fr_full_mod <- lmer(log(mean_firing_rate_vr) ~ stable_cat + (1 | mouse / session_id), data = cells_data)

summary(fr_full_mod)

fr_null_mod <- lmer(log(mean_firing_rate_vr) ~ (1 | mouse / session_id), data = cells_data)

summary(fr_null_mod)

anova(fr_full_mod, fr_null_mod)
```


# Use emmeans package to evaluate marginal means from the fit model
```{r}
fr_emm <- emmeans(fr_full_mod, "stable_cat")

pairs(fr_emm)

plot(fr_emm)
```


## ML_SNRs
# fit model and compare to a null model
Choose a gamma family as the data is continuous and > 0.
```{r}
SNRs_full_mod <- glmer(ML_SNRs ~ stable_cat + (1 | mouse / session_id), data = cells_data, family = Gamma)

summary(SNRs_full_mod)

SNRs_null_mod <- glmer(ML_SNRs ~ (1 | mouse / session_id), data = cells_data, family = Gamma)

summary(SNRs_null_mod)

anova(SNRs_full_mod, SNRs_null_mod)
```

Or use log transformed data.linear model.
```{r}
SNRs_full_mod <- lmer(ML_SNRs ~ stable_cat + (1 | mouse / session_id), data = cells_data)

summary(SNRs_full_mod)

SNRs_null_mod <- lmer(ML_SNRs ~ (1 | mouse / session_id), data = cells_data)

summary(SNRs_null_mod)

anova(SNRs_full_mod, SNRs_null_mod)
```

# Use emmeans package to evaluate marginal means from the fit model
```{r}
SNRs_emm <- emmeans(SNRs_full_mod, "stable_cat")

pairs(SNRs_emm)

plot(SNRs_emm)

```



## ML_peak_width
# fit model and compare to a null model
Choose a gamma family as the data is continuous and > 0.
```{r}
pw_full_mod <- glmer(ML_peak_width ~ stable_cat + (1 | mouse / session_id), data = cells_data, family = Gamma)

summary(pw_full_mod)

pw_null_mod <- glmer(ML_peak_width ~ (1 | mouse / session_id), data = cells_data, family = Gamma)

summary(pw_null_mod)

anova(pw_full_mod, pw_null_mod)
```

Or use log transformed data.
```{r}
pw_full_mod <- lmer(log(ML_peak_width) ~ stable_cat + (1 | mouse / session_id), data = cells_data)

summary(pw_full_mod)

pw_null_mod <- lmer(log(ML_peak_width) ~ (1 | mouse / session_id), data = cells_data)

summary(pw_null_mod)

anova(pw_full_mod, pw_null_mod)
```

# Use emmeans package to evaluate marginal means from the fit model
```{r}
pw_emm <- emmeans(pw_full_mod, "stable_cat")

pairs(pw_emm)

plot(pw_emm)

```





