---
title: "Compare firing properties across session-level firing modes"
author: "Matt Nolan"
date: "01/11/2023"
output: html_document
---

# Setup
```{r}
library(tidyverse)
library(lme4)
library(emmeans)
```

# load data
```{r}
cells_data <- readRDS(file = "cells_df_singular_values.Rda")
cells_data <- filter(cells_data, grid_cell == "TRUE")
cells_data$Lomb_classifier_ <- as_factor(cells_data$Lomb_classifier_)
```

# look at data
Replicate Figure 2F. 

Firing rate
```{r}
ggplot(cells_data, aes(Lomb_classifier_, mean_firing_rate_vr)) +
    geom_boxplot()
```

Spatial information (bits / s)
```{r}
ggplot(cells_data, aes(Lomb_classifier_, spatial_information_score_Isec_vr)) +
    geom_boxplot()
```

Spatial information (bits / s)
Distribution is more symmetrical with log transformation
```{r}
ggplot(cells_data, aes(Lomb_classifier_, log(spatial_information_score_Isec_vr))) +
    geom_boxplot()
```

Spatial information (bits / spike)
This is less informative as doesn't account for differences in firing rate.
```{r}
ggplot(cells_data, aes(Lomb_classifier_, spatial_information_score_Ispike_vr)) +
    geom_boxplot()
```

And then look at distributions
```{r}
plot_hist_groups <- function(df, plot_var) {
    plot_var <- sym(plot_var)
    ggplot(df, aes(!!plot_var)) +
        geom_histogram(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) +
        facet_wrap(vars(Lomb_classifier_))
}

plot_hist_groups(cells_data, "mean_firing_rate_vr")
```

```{r}
plot_hist_groups(cells_data, "spatial_information_score_Isec_vr")
```

```{r}
plot_hist_groups(cells_data, "ML_Freqs")
```

```{r}
plot_hist_groups(cells_data, "ML_SNRs")
```

```{r}
plot_hist_groups(cells_data, "ML_peak_width")
```

# Evaluate normality
log transformation improves fit
```{r}
qqnorm(log(cells_data$mean_firing_rate_vr), pch = 1, frame = FALSE)
qqline(log(cells_data$mean_firing_rate_vr), col = "steelblue", lwd = 2)
```

```{r}
qqnorm(log(cells_data$spatial_information_score_Isec_vr), pch = 1, frame = FALSE)
qqline(log(cells_data$spatial_information_score_Isec_vr), col = "steelblue", lwd = 2)
```

```{r}
qqnorm(cells_data$ML_SNRs, pch = 1, frame = FALSE)
qqline(cells_data$ML_SNRs, col = "steelblue", lwd = 2)
```

```{r}
qqnorm(log(cells_data$ML_peak_width), pch = 1, frame = FALSE)
qqline(log(cells_data$ML_peak_width), col = "steelblue", lwd = 2)
```


And some summary statistics
```{r}
cells_data %>% group_by(Lomb_classifier_) %>% reframe(mean = mean(spatial_information_score_Isec_vr), stdev = sd(spatial_information_score_Isec_vr))

cells_data %>% group_by(Lomb_classifier_) %>% reframe(mean = mean(mean_firing_rate_vr), stdev = sd(mean_firing_rate_vr))

cells_data %>% group_by(Lomb_classifier_) %>% reframe(mean = mean(ML_SNRs), stdev = sd(ML_SNRs))
```




## Spatial information

# fit model and compare to a null model
Choose a gamma family as the data is continuous and > 0.
```{r}
si_full_mod <- glmer(spatial_information_score_Isec_vr ~ Lomb_classifier_ + (1 | mouse / session_id), data = cells_data, family = Gamma)

summary(si_full_mod)

si_null_mod <- glmer(spatial_information_score_Isec_vr ~ (1 | mouse / session_id), data = cells_data, family = Gamma)

summary(si_null_mod)

anova(si_full_mod, si_null_mod)
```

Or fit log transformed data with linear mixed effect model.
```{r}
si_full_mod <- lmer(log(spatial_information_score_Isec_vr) ~ Lomb_classifier_ + (1 | mouse / session_id), data = cells_data)

summary(si_full_mod)

si_null_mod <- lmer(log(spatial_information_score_Isec_vr) ~ (1 | mouse / session_id), data = cells_data)

summary(si_null_mod)

anova(si_full_mod, si_null_mod)
```

```{r}
# plot(resid(si_full_mod))
qqnorm(resid(si_full_mod))
```


# Use emmeans package to evaluate marginal means from the fit model
```{r}
si_emm <- emmeans(si_full_mod, "Lomb_classifier_")

pairs(si_emm)

plot(si_emm)
```




## Spike rate
# fit model and compare to a null model
Could choose a gamma family as the data is continuous and > 0
```{r}
fr_full_mod <- glmer(mean_firing_rate_vr ~ Lomb_classifier_ + (1 | mouse / session_id), data = cells_data, family = Gamma)

summary(fr_full_mod)

fr_null_mod <- glmer(mean_firing_rate_vr ~ (1 | mouse / session_id), data = cells_data, family = Gamma)

summary(fr_null_mod)

anova(fr_full_mod, fr_null_mod)
```

Or instead fit linear mixed effect model to log transformed values
```{r}
fr_full_mod <- lmer(log(mean_firing_rate_vr) ~ Lomb_classifier_ + (1 | mouse / session_id), data = cells_data)

summary(fr_full_mod)

fr_null_mod <- lmer(log(mean_firing_rate_vr) ~ (1 | mouse / session_id), data = cells_data)

summary(fr_null_mod)

anova(fr_full_mod, fr_null_mod)
```

```{r}
#plot(resid(fr_full_mod))
#hist(resid(fr_full_mod))
qqnorm(resid(fr_full_mod))
```

# Use emmeans package to evaluate marginal means from the fit model
```{r}
fr_emm <- emmeans(fr_full_mod, "Lomb_classifier_")

pairs(fr_emm)

plot(fr_emm)
```


## ML_SNRs
# fit model and compare to a null model
Choose a gamma family as the data is continuous and > 0.
```{r}
SNRs_full_mod <- glmer(ML_SNRs ~ Lomb_classifier_ + (1 | mouse / session_id), data = cells_data, family = Gamma)

summary(SNRs_full_mod)

SNRs_null_mod <- glmer(ML_SNRs ~ (1 | mouse / session_id), data = cells_data, family = Gamma)

summary(SNRs_null_mod)

anova(SNRs_full_mod, SNRs_null_mod)
```

Or use linear model (reasonably consistent with Q-Q plot).
```{r}
SNRs_full_mod <- lmer(ML_SNRs ~ Lomb_classifier_ + (1 | mouse / session_id), data = cells_data)

summary(SNRs_full_mod)

SNRs_null_mod <- lmer(ML_SNRs ~ (1 | mouse / session_id), data = cells_data)

summary(SNRs_null_mod)

anova(SNRs_full_mod, SNRs_null_mod)
```


```{r}
# plot(resid(SNRa_full_mod))
qqnorm(resid(SNRs_full_mod))
```


# Use emmeans package to evaluate marginal means from the fit model
```{r}
SNRs_emm <- emmeans(SNRs_full_mod, "Lomb_classifier_")

pairs(SNRs_emm)

plot(SNRs_emm)

```



## ML_peak_width
# fit model and compare to a null model
Choose a gamma family as the data is continuous and > 0.
```{r}
pw_full_mod <- glmer(ML_peak_width ~ Lomb_classifier_ + (1 | mouse / session_id), data = cells_data, family = Gamma)

summary(pw_full_mod)

pw_null_mod <- glmer(ML_peak_width ~ (1 | mouse / session_id), data = cells_data, family = Gamma)

summary(pw_null_mod)

anova(pw_full_mod, pw_null_mod)
```

Or use log transformed data.
```{r}
pw_full_mod <- lmer(log(ML_peak_width) ~ Lomb_classifier_ + (1 | mouse / session_id), data = cells_data)

summary(pw_full_mod)

pw_null_mod <- lmer(log(ML_peak_width) ~ (1 | mouse / session_id), data = cells_data)

summary(pw_null_mod)

anova(pw_full_mod, pw_null_mod)
```


```{r}
# plot(resid(pw_full_mod))
qqnorm(resid(pw_full_mod))
```


# Use emmeans package to evaluate marginal means from the fit model
```{r}
pw_emm <- emmeans(pw_full_mod, "Lomb_classifier_")

pairs(pw_emm)

plot(pw_emm)

```




